---
title: "Lab 3"
author: "Ryan Morgan"
date: "February 16, 2017"
output: pdf_document
---

```{r}
library(tidyverse)

readxl:: excel_sheets("Spreadsheets/FileOne.xlsx")

readxl::read_excel("Spreadsheets/FileOne.xlsx", sheet=1) %>% glimpse()


```

First we read in the 8 total sheets.

```{r}
S11_f1 <- readxl::read_excel("Spreadsheets/FileOne.xlsx", sheet=1)

F11_f1 <-readxl::read_excel("Spreadsheets/FileOne.xlsx", sheet=2)

S12_f1 <- readxl::read_excel("Spreadsheets/FileOne.xlsx", sheet=3)

F12_f1 <- readxl::read_excel("Spreadsheets/FileOne.xlsx", sheet=4) 

S11_f2 <- readxl::read_excel("Spreadsheets/FileTwo.xlsx", sheet=1)

F11_f2 <-readxl::read_excel("Spreadsheets/FileTwo.xlsx", sheet=2)

S12_f2 <- readxl::read_excel("Spreadsheets/FileTwo.xlsx", sheet=3)

F12_f2 <- readxl::read_excel("Spreadsheets/FileTwo.xlsx", sheet=4) 

```

Add column for what semester it is.




Start by making sure columns are labeled.

```{r}
colnames(F11_f1)[c(1,2)]=c("ID","Test")
colnames(F11_f2)[c(2)]=c("Test")
colnames(S12_f1)[c(1,2)]=c("ID","Test")
colnames(S12_f2)[c(2)]=c("Test")

colnames(F12_f1)[c(1,2)]=c("ID","Test")
colnames(F12_f2)[c(2)]=c("Test")
colnames(S11_f1)[c(1,2)]=c("ID","Test")
colnames(S11_f2)[c(2)]=c("Test")

```

Second semester file 1 has an empty column. So remove that.

```{r}
F11_f1<- F11_f1[,-15]
```
First semester, file 2 missing math3. So go back and add a column of "NA" values. 

```{r}
S11_f2$"MTH 3" <- rep(NA, nrow(S11_f2))
```

Last semester, file 2 the id's are too high by 1000.

```{r}
F12_f2$ID = F12_f2$ID - 1000
```

Now combine the sheets back together.


File1 <- rbind(S11_f1, F11_f1, S12_f1,F12_f1)

File2 <- rbind(S11_f2, F11_f2, S12_f2,F12_f2)

Remove the extra space after treatment 2 in the file 1.

```{r}
colnames(File1)[18]="Treatment Part 2"
```

Some test titles are all caps (ex: PRE) while some are not capitilized (ex: Pre). 

```{r}
File1$Test <- gsub("PRE", "Pre", File1$Test)

File1$Test <- gsub("POST", "Post", File1$Test)

File2$Test <- gsub("PRE", "Pre", File2$Test)

File2$Test <- gsub("POST", "Post", File2$Test)

```

Extract the id, gender, and charachteristic from each file. 


newnew doesn't have any NA values (hopefully)
```{r}
idgenchar1 <- File1[,c(1,15,16)]

idgenchar2 <- File2[,c(1,36,37)]

idgenchar <- unique(rbind(idgenchar1,idgenchar2))


new <- idgenchar[-which(is.na(idgenchar$GENDER)),]

newnew <- new[-which(is.na(new$Characteristic)),]

```

Remove the gender and char columns from other files. 
```{r}
File1 <- File1[,-c(15,16)]

File2 <- File2[,-c(36,37)]

```

```{r}

final <- left_join(newnew,File1,by=c("ID"))

done <- left_join(final, File2, by=c("ID","Test"))

test <- done %>% mutate(Semester= floor(ID/10000))

test$Semester = gsub(0,1,test$Semester)

```



After merging two data frames, we can bring in the third one which has gender and charachteristic.


We would want each student to have just 1 row.


```{r}
testing <- test %>% mutate(TotalScore = Total.y + Total.x) %>% select(ID, GENDER, TotalScore, Semester, Test) 


pleaseWork <- unique(testing)

pleaseWork%>% spread(key=Test, value=TotalScore) %>% ggplot(aes(x=Pre,y=Post, colour=GENDER)) + geom_point()+facet_wrap(~Semester)



```
